An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-07-10 EconGrowth.Rnw
CD model now uses resampling information for confidence intervals.
Added notes about elasticity of substitution.
Edited the to-do list to reflect fact that we're now using resampled data to calculate SF confidence intervals.
Now including CES resampling graphs.

\section*{To-do List}
\begin{itemize}
\item{Heun: Add option for $m = 1$ in single-factor models.}
\item{Heun: Figure out why CESe for ZA with $x$ fails during resampling.}
\item{Heun: Review all assumptions for CES with energy models. 
There are several assumptions for rho and rho1 for various countries. 
When we are resampling, we probably don't want to specify those values.}
\item{Heun: Re-implement confidence interval calculations to use
resampling methods.}
      \begin{itemize}
      \item{SF}
      \item{CD}
      \item{\st{CDe}}
      \item{CES}
      \item{CESe}
      \item{LINEX}
      \end{itemize}
\item{Heun: Include cloud graphs for model parameters 
estimated by resampling methods.}
      \begin{itemize}
      \item{\st{SF}}
      \item{CD}
      \item{\st{CDe}}
      \item{CES}
      \item{CESe}
      \item{LINEX}
      \end{itemize}
\item{\st{Heun: Decide if the single factor models should include constraints
on $\alpha$, $\beta$, and $\gamma$. At present, the single-factor model allows
values outside the ranges $\alpha \in (0,1)$, $\beta \in (0,1)$, and $\gamma \in (0,1)$.}}
\item{\st{Heun: Make smaller cloud plots for resample coefficients for those 
countries where $u$ is involved.}}
\item{Heun: Investigate possible conflicting use of exergy by Warr and Ayres.}
\item{Heun: Get AIC calculations working again.}
\item{Heun: Fix the energy sankey diagram from the seminar}
\item{Heun: Include adjusted sankey diagram in the seminar}
\item{Heun: Add CES production function fits}
      \begin{itemize}
      \item{Write up the decisions about parameter estimation for the 
            CES production function fits}
      \item{Revisit my assumptions for the CES production function 
            (with energy) for each country.}
      \item{Fix the problems with $\sigma_1$ and $\sigma$ values in 
            the CES parameter tables in the appendix}
      \item{Add CES rows to the AIC table. Code is in the 
            \texttt{createAICTable} function, but it is not yet working. 
            Need to find a way to calculate AIC for the object that is 
            returned from the \texttt{cesEst} function.}
      \end{itemize}
\item{Heun: Fix warnings of the form \texttt{Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562} in the code that generates the Cobb-Douglas with energy fits.}
\item{de Wit: review, critique, and contextualize.}
\end{itemize}

@@

\noindent Equation~\ref{eq:CES} is a CES production function with 
capital stock ($k$) and labor ($l$) factors of production. 
Equation~\ref{eq:CESe} augments Equation~\ref{eq:CES} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CES} is a degenerate form of Equation~\ref{eq:CESe} 
where $\delta = 1$. The elasticity of substitution between capital ($k$) 
and labor ($l$) is given by $\sigma_1 = \frac{1}{1+\rho_1}$, 
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$, $\delta \in [0, 1]$ describes the importance 
of ($kl$) relative to ($e$), and $\gamma$ is a fitting parameter, 
expected to be close to unity.

@@

@@

The CES model without and with energy is given by 
Equations~\ref{eq:CES} and \ref{eq:CESe}, respectively. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\gamma$, $\lambda$, $\delta_1$, $\rho_1$, $\delta$, and $\rho$. 
**** More details here about which algorithms were used, etc. 
Complete after our work this summer. **** 
Elasticities of substitution $\sigma_1$ and $\sigma$ 
(see Section~\ref{sec:CES}) were estimated by the delta method 
(**** is this correct? Ask Pruim. ****).

@@

@@

>>>>

\section*{To-do List}
\begin{itemize}
\item{Heun: Add option for $m = 1$ in single-factor models.}
\item{Heun: Figure out why CESe for ZA with $x$ fails during resampling.}
\item{Heun: Review all assumptions for CES with energy models. 
There are several assumptions for rho and rho1 for various countries. 
When we are resampling, we probably don't want to specify those values.}
\item{Heun: Re-implement confidence interval calculations to use
resampling methods.}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{\st{CES}}
      \item{\st{CESe}}
      \item{LINEX}
      \end{itemize}
\item{Heun: Include cloud graphs for model parameters 
estimated by resampling methods.}
      \begin{itemize}
      \item{\st{SF}}
      \item{\st{CD}}
      \item{\st{CDe}}
      \item{CES}
      \item{CESe}
      \item{LINEX}
      \end{itemize}
\item{\st{Heun: Decide if the single factor models should include constraints
on $\alpha$, $\beta$, and $\gamma$. At present, the single-factor model allows
values outside the ranges $\alpha \in (0,1)$, $\beta \in (0,1)$, and $\gamma \in (0,1)$.}}
\item{\st{Heun: Make smaller cloud plots for resample coefficients for those 
countries where $u$ is involved.}}
\item{Heun: Investigate possible conflicting use of exergy by Warr and Ayres.}
\item{Heun: Get AIC calculations working again.}
\item{Heun: Fix the energy sankey diagram from the seminar}
\item{Heun: Include adjusted sankey diagram in the seminar}
\item{Heun: Add CES production function fits}
      \begin{itemize}
      \item{Write up the decisions about parameter estimation for the 
            CES production function fits}
      \item{Revisit my assumptions for the CES production function 
            (with energy) for each country.}
      \item{Fix the problems with $\sigma_1$ and $\sigma$ values in 
            the CES parameter tables in the appendix}
      \item{Add CES rows to the AIC table. Code is in the 
            \texttt{createAICTable} function, but it is not yet working. 
            Need to find a way to calculate AIC for the object that is 
            returned from the \texttt{cesEst} function.}
      \end{itemize}
\item{Heun: Fix warnings of the form \texttt{Warning:  step factor 0.000488281 reduced below ’minFactor’ of 0.000976562} in the code that generates the Cobb-Douglas with energy fits.}
\item{de Wit: review, critique, and contextualize.}
\end{itemize}

@@

\noindent Equation~\ref{eq:CES} is a CES production function with 
capital stock ($k$) and labor ($l$) factors of production. 
Equation~\ref{eq:CESe} augments Equation~\ref{eq:CES} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CES} is a degenerate form of Equation~\ref{eq:CESe} 
where $\delta = 1$. The elasticity of substitution between capital ($k$) 
and labor ($l$) is given by $\sigma_1 = \frac{1}{1+\rho_1}$, 
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$, $\delta \in [0, 1]$ describes the importance 
of ($kl$) relative to ($e$), and $\gamma$ is a fitting parameter, 
expected to be close to unity. As $\sigma \rightarrow \infty$, 
($kl$) and ($e$) are perfect substitutes. As $\sigma \rightarrow 0$, 
($kl$) and ($e$) are perfect complements: no substitution is possible.

@@

Converting the resample data into 95\% confidence intervls 
provides data for Figure~\ref{fig:CD_Params_Graph}.

@@

The CES model without and with energy is given by 
Equations~\ref{eq:CES} and \ref{eq:CESe}, respectively. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\gamma$, $\lambda$, $\delta_1$, $\rho_1$, $\delta$, and $\rho$. 
**** More details here about which algorithms were used, etc. 
Complete after our work this summer. ****

@@

Coefficients for each resampled CES model (without energy)
are shown in Figures~\ref{fig:CES_lambda_gamma_Cloud_Graph} 
and \ref{fig:CES_sigma_1_delta_1_Cloud_Graph}.

<<CES_lambda_gamma_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (without energy) resample model parameters $\\lambda$ and $\\gamma$.", eval=TRUE>>=
cesLambdaGammaResamplePlot(energyType=NA)
@

<<CES_sigma_1_delta_1_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (without energy) resample model parameters $\\sigma_1$ and $\\delta_1$.", eval=TRUE>>=
cesSigma_1Delta_1ResamplePlot(energyType=NA)
@

Coefficients for each resampled CES model (with $q$)
are shown in 
Figures~\ref{fig:CESq_lambda_gamma_Cloud_Graph}--\ref{fig:CESq_sigma_delta_Cloud_Graph}.

<<CESq_lambda_gamma_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (with $q$) resample model parameters $\\lambda$ and $\\gamma$.", eval=TRUE>>=
cesLambdaGammaResamplePlot(energyType="Q")
@

<<CESq_sigma_1_delta_1_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (with $q$) resample model parameters $\\sigma_1$ and $\\delta_1$.", eval=TRUE>>=
cesSigma_1Delta_1ResamplePlot(energyType="Q")
@

<<CESq_sigma_delta_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (with $q$) resample model parameters $\\sigma$ and $\\delta$.", eval=TRUE>>=
cesSigmaDeltaResamplePlot(energyType="Q")
@

Coefficients for each resampled CES model (with $x$)
are shown in 
Figures~\ref{fig:CESx_lambda_gamma_Cloud_Graph}--\ref{fig:CESx_sigma_delta_Cloud_Graph}.

<<CESx_lambda_gamma_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (with $x$) resample model parameters $\\lambda$ and $\\gamma$.", eval=TRUE>>=
cesLambdaGammaResamplePlot(energyType="X")
@

<<CESx_sigma_1_delta_1_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (with $x$) resample model parameters $\\sigma_1$ and $\\delta_1$.", eval=TRUE>>=
cesSigma_1Delta_1ResamplePlot(energyType="X")
@

<<CESx_sigma_delta_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES (with $x$) resample model parameters $\\sigma$ and $\\delta$.", eval=TRUE>>=
cesSigmaDeltaResamplePlot(energyType="X")
@

Coefficients for each resampled CES model (with $u$)
are shown in 
Figures~\ref{fig:CESu_lambda_gamma_Cloud_Graph}--\ref{fig:CESu_sigma_delta_Cloud_Graph}.

<<CESu_lambda_gamma_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=threePanelGraphHeight, fig.cap="CES (with $u$) resample model parameters $\\lambda$ and $\\gamma$.", eval=TRUE>>=
cesLambdaGammaResamplePlot(energyType="U")
@

<<CESu_sigma_1_delta_1_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=threePanelGraphHeight, fig.cap="CES (with $u$) resample model parameters $\\sigma_1$ and $\\delta_1$.", eval=TRUE>>=
cesSigma_1Delta_1ResamplePlot(energyType="U")
@

<<CESu_sigma_delta_Cloud_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=threePanelGraphHeight, fig.cap="CES (with $u$) resample model parameters $\\sigma$ and $\\delta$.", eval=TRUE>>=
cesSigmaDeltaResamplePlot(energyType="U")
@

@@

Converting the resampling data to 95\% confidence intervals
provides results shown in 
Figures~\ref{fig:SFk_Params_Graph}--\ref{fig:SFu_Params_Graph}.
